Relative Stable Maps in Genus One via Central Alignments

### 2018-12-30 GenusOneRelativeMaps.tex
Finished second example

\draw [->] (2,2) -- (3,2);
\draw (3,2) node[above]{$x_1$};

@@

\begin{example}
Consider now an element of $\Mcal$ whose associated tropical map has the corresponding combinatorial type:

>>>>

\draw [->] (2,2) -- (3,2);
\draw (3,2) node[above]{$x_1$};
\draw (2.5,2) node[below,red]{$3$};

@@

\begin{example}
Consider now an element of $\Mcal$ whose associated tropical map has the corresponding combinatorial type:
\begin{center}
\begin{tikzpicture}
\draw [fill] (0,4) circle[radius=2pt];
\draw (0,4) node[left]{$\mathsq{C}_1$};
\draw (0,4) -- (2,3);
\draw (1,3.5) node[above,color=red]{$1$};
\draw (1,3.5) node[below]{$e_1$};

\draw [->] (2,3) -- (3,3);
\draw (3,3) node[above]{$x_1$};
\draw (2.5,3) node[above,red]{$1$};

\draw (2,3) -- (2,2);
\draw [fill] (2,2) circle[radius=2pt];
\draw (2,2.5) node[right]{$e_3$};
\draw (2.075,2.5) node[left,red]{$0$};
\draw (2,2) node[below]{$\mathsq{C}_3$};

\draw [->] (2,2) -- (3,2);
\draw (3,2) node[above]{$x_2$};
\draw (2.5,2) node[below,red]{$2$};

\draw [fill=white] (2,3) circle[radius=2pt];
\draw (2,3) node[above]{$\mathsq{C}_0$};

\draw [->,color=red] (2,1.25) -- (2,0.5);

\draw [->,color=red] (0,0) -- (3,0);
\draw (0,0) [fill=red,color=red] circle[radius=2pt];
\draw (3,0) node[right,color=red]{$\Sigma(\PP^N|H)$};
\end{tikzpicture}
\end{center}
As before, expansion factors are indicated in red and the edge lengths are $e_1,e_2,e_3$. The tropical moduli space is $\RR_{\geq 0}^3$ generated by these three lengths, subject to the continuity condition $e_1=2e_2$. Thus the moduli space is isomorphic to $\RR_{\geq 0}^2$ generated by $e_2$ and $e_3$. In order to have an alignment, at least one of $\lambda(\mathsq{C}_1)$ and $\lambda(\mathsq{C}_2)$ must be equal to the radius $\delta$. Note that
\begin{align*} \lambda(\mathsq{C}_1) & = e_1 = 2e_2 \\
\lambda(\mathsq{C}_2) & = e_2 + e_3\end{align*}
and without more information we cannot say which of these is larger. The subdivision of $(\RR_{\geq 0})^2$ is obtained by dividing the cone into regions where different order relations hold amongst the distances $\lambda(\mathsq{C}_1),\lambda(\mathsq{C}_2),\lambda(\mathsq{C}_3)$. The walls of this subdivison correspond to where some of these distances are equal. Note that in this setting we always have $\lambda(\mathsq{C}_2) > \lambda(\mathsq{C}_3)$ (at least, as long as we remain in the interior of the cone). The remaining possibilities are:
\begin{align*} \lambda(\mathsq{C}_1) & = \lambda(\mathsq{C}_3) \qquad (\Leftrightarrow e_1 = e_3 \Leftrightarrow 2e_2 = e_3) \\
\lambda(\mathsq{C}_1) & = \lambda(\mathsq{C}_2) \qquad (\Leftrightarrow e_1 = e_2 + e_3 \Leftrightarrow e_2 = e_3). \end{align*}
Thus, the subdivision of the tropical moduli space $(\RR_{\geq 0}^2)_{e_2 e_3}$ is given by:
\begin{center}
\begin{tikzpicture}[scale=1]
% Right-hand box
\fill [fill=gray!30!white] (5,0) -- (8.75,0) -- (8.75,3.75) -- (5,3.75) -- cycle;
\draw [fill] (5,0) circle[radius=2pt];
\draw [->] (5,0) -- (9,0);
\draw (9,0) node[right]{$e_2$};
\draw [->] (5,0) -- (5,4);
\draw (5,4) node[above]{$e_3$};

% Left-hand box
\fill [fill=gray!30!white] (-3,0) -- (0.75,0) -- (0.75,3.75) -- cycle;
\fill [fill=gray!50!white] (-3,0) -- (0.75,3.75) -- (-1.125,3.75) -- cycle;
\fill [fill=gray!30!white] (-3,0) -- (-1.125,3.75) -- (-3,3.75) -- cycle;
\draw [fill] (-3,0) circle[radius=2pt];
\draw [->] (-3,0) -- (1,0);
\draw (1,0) node[right]{$e_2$};
\draw [->] (-3,0) -- (-3,4);
\draw (-3,4) node[above]{$e_3$};

% 2e_2=e_3 line
\draw [->] (-3,0) -- (-1,4);
\draw (-1,4) node[above]{$2e_2=e_3$};

% e_2=e_3 line
\draw [->] (-3,0) -- (1,4);
\draw (1,4) node[above]{$e_2=e_3$};

% arrow
\draw [->] (2,2) -- (4,2);
\end{tikzpicture}
\end{center}
The cones of the subdivison index the logarithmic strata in a neighbourhood of the preimage in $\VZ$ of our chosen element of $\Mcal$. These can be described as follows:
\begin{center}
\begin{tikzpicture}[scale=1.8]
% Left-hand box
\fill [fill=gray!30!white] (-3,0) -- (0.75,0) -- (0.75,3.75) -- cycle;
\fill [fill=gray!50!white] (-3,0) -- (0.75,3.75) -- (-1.125,3.75) -- cycle;
\fill [fill=gray!30!white] (-3,0) -- (-1.125,3.75) -- (-3,3.75) -- cycle;
\draw [fill] (-3,0) circle[radius=1pt];
\draw (-3,0) -- (1,0);
\draw (1,0) node[right,red]{$e_3=0$};
\draw (-3,0) -- (-3,4);
\draw (-3,4) node[above,red]{$e_1=e_2=0$};

% 2e_2=e_3 line
\draw (-3,0) -- (-1,4);
\draw (-1,4) node[above,red]{$\lambda(\mathsq{C}_1)=\lambda(\mathsq{C}_3)$};

% e_2=e_3 line
\draw (-3,0) -- (1,4);
\draw (1,4) node[above,red]{$\lambda(\mathsq{C}_1)=\lambda(\mathsq{C}_2)$};

\draw (0,1.5) node[below,red]{$\lambda(\mathsq{C}_1) > \lambda(\mathsq{C}_2)$};

\draw (-0.6,3.4) node[below,red]{$\lambda(\mathsq{C}_3) < \lambda(\mathsq{C}_1)$};
\draw (-0.6,3.1) node[below,red]{$< \lambda(\mathsq{C}_2)$};

\draw (-2.2,3.4) node[below,red]{$\lambda(\mathsq{C}_1) < \lambda(\mathsq{C}_3)$};
\end{tikzpicture}
\end{center}
Geometrically, there are four codimension--$1$ logarithmic strata of $\VZ$ intersecting our chosen neighbourhood, corresponding to the rays in the above picture. Two of these -- those corresponding to $\{ e_1=e_2=0 \}$ and $\{ e_3=0 \}$ -- are the proper transforms of codimension--$1$ strata in $\Mcal$. Here some of the edge lengths are set to zero, and the corresponding logarithmic strata consist of  stable maps where the relevant nodes are smoothed. Notice that although the curve has three nodes, there are only two such strata: the nodes $q_1$ and $q_2$ cannot be smoothed independently because of the relation $e_1=2e_2$.

The remaining two codimension--$1$ strata correspond to where certain of the vertex distances become equal. In these strata, none of the nodes are smoothed. From the construction of the subdvision, we see that they both map onto the codimension--$2$ stratum of $\Mcal$ in which all of the nodes persist; this coheres with the fact that they should be thought of as exceptional loci of the blow-up. The extra dimension of moduli comes from the choice of alignment.

There are three codimension--$2$ strata corresponding to different strict orderings of the vertex distances. Note that the strata corresponding to $\{e_1=e_2=0\}$ and $\{e_3=0\}$, which intersected in $\Mcal$, no longer intersect in $\VZ$. The picture is something like:
\begin{center}
\begin{tikzpicture}[scale=0.5]
\draw (-6.5,0.5) ellipse (4 and 2);
\draw (-6.5,-1.5) node[below]{$\Mcal$};
\draw [very thick] plot [smooth] coordinates{(-10,0.25) (-8,0.75) (-6,0)};
\draw [very thick] plot [smooth] coordinates{(-7,0) (-5,0.75) (-3,0.5)};

\draw (-21.5,0.5) ellipse (7 and 2);
\draw (-21.5,-1.5) node[below]{$\VZ$};
\draw [very thick] plot [smooth] coordinates{(-28,0.25) (-26,0.75) (-24,0)};
\draw plot [smooth] coordinates{(-25,0) (-23,0.75) (-21,0)};
\draw plot [smooth] coordinates{(-23,0) (-20,0.75) (-18,0)};
\draw [very thick] plot [smooth] coordinates{(-20,0) (-17,0.75) (-15,0)};

\draw [->] (-14,0.5) -- (-11,0.5);
\end{tikzpicture}
\end{center}
